US General Society Survey Analysis

### 2024-03-17 paper/paper.qmd
.

# Models/Results

@@

### Voters in 2016

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-hillary2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"

# only choose values "Hillary Clinton" and "Donald Trump"
ces2020_votes <- ces2020 %>%
  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump", "Other", "Did not vote"))

@@

plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)") +
  theme_minimal()
plot_hillary

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"

#Trum Plot
plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
  theme_minimal()
plot_trump
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-other2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dnv2016
#| fig-cap: "Linear Regression comparing Percentage of Non-Voters and Year of Birth (2016)"

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-biden2020
#| fig-cap: "Linear Regression comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"

@@

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2020
#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump Year of Birth (2020)"

@@

plot_trump_2
```

@@

@@

  It is near impossible to have data sets that are completely error-free and the same is true for this data-set. The data-set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for. A lot of marginalized opinions are being left out when we clean the data and we only look for people that could vote. For example refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care. 
  Another weakness in our data-set is what occurs with a lot of data-sets, which is missing values. This data set also contained them and those needed to be cleaned so that our analysis would not sway. Speaking of missing values, insignificant values were ignored in some of the models as well. For, example some of the presidential candidates got so little votes that including them would have actually hindered our ability to see the trends so we had to group them in others. This was beneficial for this paper but  if we do indeed continue to analyse this paper for other purposes, this might cause problems.

  In terms of next steps for this paper, we look forward to predict the 2024 election results by comparing the what the general priorities of the population is. We could look at population by immigration status, gender, age, or their previous 2020 voting results.  If we are able to correctly predict the results based on the model continuously, it would help understand the priorities of american citizens; knowledge that would in turn help us help them.

>>>>

# Models

@@

### Aging Voters (2016)

In the following graphs we will be looking at the percentage of voters from each birth year that voted for Hillary Clinton, Donald Trump,and other candidates in the 2016 US Presidential election, we will also take a look at those who did not vote.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-hillary2016
#| fig-cap: "Comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"

ces2020_votes <- ces2020 %>%
  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump", "Other", "Did not vote"))

@@

#R value
hillary_data <- percentage_data_long %>% filter(Candidate == "HillaryClinton")

R_hillary <- cor(hillary_data$birthyr, hillary_data$Percentage)

plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()
plot_hillary

@@

In @fig-hillary2016 we can see that the younger the voters, the smaller percentage of their birth year seem to have voted for Hillary Clinton in the 2016 presidential election. This was rather surprising for us, as we expected young voters more likely to vote for Hillary Clinton and the Democratic Party, as younger people are often more progressive than their older counterparts. @gig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2016
#| fig-cap: "Comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"

trump_data <- percentage_data_long %>% filter(Candidate == "DonaldTrump")

R_trump <- cor(trump_data$birthyr, trump_data$Percentage)

#Trum Plot
plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()
plot_trump
```

While we look at the graph for Donald Trump in the 2016 election at first glance it does seem that the younger you are the more likey you were to have voted for trump in the 2016 election, but if we dig a bit deeper we can see that there seem to be several outliers on this graph, and the data points seem to somewhat be bouncing around a bit. This is confirmed when we look at our correlation coefficient $ R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-other2016
#| fig-cap: "Comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"

other_data <- percentage_data_long %>% filter(Candidate == "Other")

R_other <- cor(other_data$birthyr, other_data$Percentage)

@@

@fig-other2016 does not seem to give us to much information about the age of those who voted for other candidates in the 2016 presidential elections, such as Gary Johnson and Jill Stein. While all these candidates received a combined less than 5% of the total vote, and less than 5% of the share of the vote from each birth year, we believe that transparency is incredibly important when doing data heavy research and while this may not seem as important as the graphs for Hillary Clinton and Donald Trump, this graph absolutely has a place in this paper.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dnv2016
#| fig-cap: "Comparing Percentage of Non-Voters and Year of Birth (2016)"

dnv_data <- percentage_data_long %>% filter(Candidate == "Did not vote")

R_dnv <- cor(dnv_data$birthyr, dnv_data$Percentage)

@@

Here we will look at the members of the population that did not vote. @fig-dnv2016

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-biden2020
#| fig-cap: "Comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"

@@

#Rvalue Biden
biden_data <- percentage_data_long %>% filter(Candidate == "Biden")

R_biden <- cor(biden_data$birthyr, biden_data$Percentage)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2020
#| fig-cap: "Comparing Percentage of Votes For Donald Trump Year of Birth (2020)"

@@

plot_trump_2

#rvalue Trump 2020

trump_2_data <- percentage_data_long %>% filter(Candidate == "Trump")

R_trump_2 <- cor(trump_2_data$birthyr, trump_2_data$Percentage)
```

@@

# Results

@@

Weaknesses and next steps should also be included.
